Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-09-28 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Edits to 3 and 4

These instrumented non-combustible wall experiments provide a data set for the fires conducted that can be compared against ignition data for gypsum wallboard from the literature to determine the potential for ignition of the paper surface of gypsum wallboard at a given distance from the wall with a given fuel. The Ignition Handbook by Babrauskas is the most comprehensive set of ignition data and related information ever written~\cite{Babrauskas:2003}.  Table~19 in the Ignition Handbook includes calculated ignition temperatures from experimental data from two separate sources.

@@

Of course the temperatures of the thermocouples in the wall are a result of the energy transferred to the wall and the thermo-physical properties of the wall material. The properties of the calcium silicate board are simlar but not identical to gypsum wallboard.  The heat flux data is independent of the wall material, so it may be a better predictor of the fire damage to gypsum wallboard when exposed to the burners at different distances.    All of the fuels generated heat fluxes in excess of 26~kW/m$^2$.  However only the burner against the wall positions provided a combination of heat flux and temperature that had the potential to ignite the paper on the gypsum wallboard based on the handbook data.  The experiments conducted with the instrumented gypsum wall board will provide data for comparsion with data provided in this chapter and provide insight in to where the gypsum wall board ignites.

@@

\begin{figure}[p]
	\centering
	\includegraphics[width=2.5in]{../Figures/Soot_pattern}
	\includegraphics[width=2.5in]{../Figures/Soot_close}	\\
	\caption[Photographs showing clear lines of demarcation with gasoline]{Photographs showing clear lines of demarcation from a fire pattern generated on painted gypsum wallboard by a gasoline fueled burner.  Burned vs unburned area is clearly defined, where as the difference between sooted and thermally discolored areas is not as clear.}
	\label{Soot_pattern}
\end{figure}

@@

Figure~\ref{IWGB_NG_patterns} provides a comparison between the area of discoloration of the paint with the natural gas burner positioned at 0.15~m away from the wall versus the area of consumed and penetrtaed paint/paper layer which exposed the gypsum underneath.

>>>>

These instrumented non-combustible wall experiments provide a data set for the fires to be conducted that can be compared against ignition data for gypsum wallboard from the literature to determine the potential for ignition of the paper surface of gypsum wallboard at a given distance from the wall with a given fuel. The Ignition Handbook by Babrauskas is the most comprehensive set of ignition data and related information ever written~\cite{Babrauskas:2003}.  Table~19 in the Ignition Handbook includes calculated ignition temperatures from experimental data from two separate sources.

@@

Of course the temperatures of the thermocouples in the wall are a result of the energy transferred to the wall and the thermo-physical properties of the wall material. The properties of the calcium silicate board are simlar but not identical to gypsum wallboard.  The heat flux data is independent of the wall material, so it may be a better predictor of the fire damage to gypsum wallboard when exposed to the burners at different distances.    All of the fuels generated heat fluxes in excess of 26~kW/m$^2$.  However only the burner against the wall positions provided a combination of heat flux and temperature that had the potential to ignite the paper on the gypsum wallboard based on the handbook data.  Based on the heat flux data from Figures~\ref{NCTW_HF_Center_Comp} the natural gas burner should burn the paint and paper off of the face of the gypsum wallboard up to a height of between 0.6~m amd 0.8~m above the burner.  Similarly, the gasoline burner should generate a burn area between 0.8~m and 1.0~m high on a wall composed of gypsum wallboard.  Using the same analysis, the polyurethane foam burner would create a shorter burn region approximately 0.4~m to 0.6~m.  The experiments conducted with the instrumented gypsum wall board will provide data for comparsion with data provided in this chapter and provide insight in to where the may gypsum wall board ignite.

@@

\begin{figure}[p]
	\centering
	\includegraphics[width=2.5in]{../Figures/Soot_pattern}
	\includegraphics[width=2.5in]{../Figures/Soot_close}	\\
	\caption[Photographs showing lines of demarcation with gasoline]{Photographs showing lines of demarcation from a fire pattern generated on painted gypsum wallboard by a gasoline fueled burner.  Burned vs unburned area is clearly defined, where as the difference between sooted and thermally discolored areas is not as clear.}
	\label{Soot_pattern}
\end{figure}

@@

Figure~\ref{IWGB_NG_patterns} provides a comparison between the area of discoloration of the paint with the natural gas burner positioned at 0.15~m away from the wall versus the area of consumed and penetrated paint/paper layer which exposed the gypsum underneath.
